---
title: "Power Play: Corruption, Constitutional Change, Leadership"
description: "Ayla Aliyeva | Fall 2023 | Gov 50 Final Project"
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    code_folding: true 
---

## Set up chunk
```{r}
library(tidyverse)
library(ggthemes)
library(wesanderson)
library(jtools)
library(modelsummary)
```



## Preparing the data

```{r}

## Reading data files

violence <- read_csv("data(final project gov 50)/Violence/violence.csv")

dpi <- read_csv("data(final project gov 50)/DPI2020/DPI2020.csv")


## Calculating overall number of constitutional changes for each country over a given period

violence <- 
  violence |>
  group_by(country) |>
  mutate(constch_overall = constchg70 + constchg80) 


## Calculating the average number of years Chief Executive has been in office for each country

colnames(dpi) [1] <- "country"

dpi <- 
   dpi |>
  group_by(country) |>
  mutate(avg_yrsoffc = mean(yrsoffc, na.rm = TRUE)) |>
  filter(yrsoffc >= 0)


## Filtering one of the data sources to only include needed decades

dpi <- 
  dpi |>
  filter(year >= 1970 & year <= 1989) 
  
  
## Merging data sources by country

dpi <- 
  dpi|>
  mutate(country = toupper(country))

 data <- 
   merge(violence, dpi, by = "country")
 
 
 ## Tidying up the data: selecting needed variables
 
 data <- data |>
   select(country,year, constchg70, constchg80, constch_overall, corrupti, yrsoffc, avg_yrsoffc)
 


```

## Introduction

Do higher levels of corruption lead to more frequent constitutional changes? If so, is there a direct correlation between this trend and the average number of years the Chief Executive of a country stays in power? In this study, I examine data from a variety of countries in order to explore the relationship between these variables throughout the two decades of the 1970s and 1980s. I hypothesise that higher levels of corruption are positively correlated with more frequent constitutional changes and a higher average period within which a Chief Executive remains in office. This hypothesis is grounded in the theory that governments with higher corruption levels are more likely to revise legislation in order to hold on to power.

 The significance of this study is in its attempt to explore the implications of corruption in the misuse of power to influence constitutional frameworks and governance. In the case that the data supports the hypothesis, this would point to a country’s level of corruption as a determinant of both the nature of its legal system, as well as the course of its executive leadership.


## Data 

 The data source used in this study is one that has been gathered by merging the Political Violence Dataset, as well as the Inter-American Development Bank “Database of Political Institutions” to ultimately extract the key dependent and independent variables. 
 
The dependent variable used in this research is the overall number of major constitutional changes for countries throughout the period of the 1970s and 1980s, derived from the two variables relaying the number of constitutional changes for a given country within each of the two decades, its measurement method being through a compilation of government records. The key independent variables are the countries’ level of corruption, as well as the average number of years the Chief Executive remains in office. The level of corruption is measured using 1980-1989 Knack and Keefer measure of corruption, with a lower number indicating a lower level of corruption. The latter independent variable relays the official number of years the Chief Executive of the time remained in office, calculated cumulatively for each Chief Executive as of January 1st of each year, derived from public records.

 The bar plot below summarises the frequency of constitutional change, a discrete variable, across countries between the 1970s and 1980s, demonstrating that a majority countries in the dataset had not gone through constitutional changes within the given period. The second highest bar signifies countries that had undergone 1 constitutional change, with the lowest bar showing countries that had undergone 2 constitutional changes.


```{r}

## Creating an object that extracts first observation for each country
countryonce_data <-
  data |>
  group_by(country) |>
  slice(1)


## Bar plot summarizing key dependent variable
countryonce_data |>
  ggplot(mapping = aes(x = constch_overall)) + 
  geom_bar(fill = "lightsteelblue", colour = "darkblue") + 
  labs(title = "Frequency of Constitutional Change Across Countries", y = "Number of Countries", x = "Number of Major Constitutional Changes (1970s-1980s)") + 
  theme_calc() +
  theme(text = element_text("Georgia"))


```






## Results


In order to explore the main relationships of interest, I constructed a scatter plot, followed by a linear regression of the results. As shown by the estimated line of best fit in the graph below, the correlation between the number of constitutional changes and the 1980-1989 Knack and Keefer index of corruption is negative. As mentioned above, a higher corruption index signifies a lower actual level of corruption. Therefore, according to the plot, higher corruption is associated with a higher number of constitutional changes.

 The linear regression further supports this relationship. The corrupti (corruption index) coefficient of -0.035 estimates that for every unit increase in the corruption index, there is a 0.035 decrease in the number of constitutional changes.  The p value being below 0.001 shows that the coefficient is statistically significant, meaning that we may reject the null hypothesis of no effect. Furthermore, the relatively small standard error (which shows how far, on average, the actual points of observation lie from the estimated line of regression) of 0.008 implies the regression to be a rather precise estimate.
 
 In regards to the relationship between the number of constitutional changes and the average number of years a Chief Executive remains in office (the second independent variable), the regression shows a positive correlation. More specifically, the coefficient of 0.001 suggests that for every unit increase in the mean Chief Executive term in office, there is a 0.001 increase in the number of constitutional changes. However, the p value of 0.115, means that at a significance level of 0.05, the coefficient is not statistically significant. Therefore, in this case, we cannot reject the null hypothesis.
 
 Considering the analysis above, while we can establish a correlation between the dependent and independent variables, we cannot be certain that the relationship between them is causal. This is due to the possibility of omitted variables as well as potential confounders, such as factors influencing the overall economic and political stability in the given countries.


```{r}

## Creating a variable with categories describing average years of countries' Chief Executive in office

data <- 
  data |>
  mutate(avgyrs_range = case_when(
    avg_yrsoffc <= 5 ~ "5 Years and Under",
    avg_yrsoffc > 5 & avg_yrsoffc <= 15 ~ "Between 5 and 15 Years",
    avg_yrsoffc > 15 & avg_yrsoffc <= 25 ~ "Between 15 and 25 Years",
    avg_yrsoffc >= 25 ~ "Over 25 Years"
  )) 


## Plot of main analysis 

fit_visual <- 
  data |>
  ggplot(mapping = aes(x = corrupti, y = constch_overall, colour = avgyrs_range)) + 
  geom_point() + 
  geom_smooth(method = lm, colour = "salmon") + 
  labs(title = "Linear Regression on Corruption and Constitutional Changes", y = "Number of Constitutional Changes", x = "Corruption Index", colour = "Mean Chief Executive Term") + 
  theme_calc() +
  scale_color_manual(values = wes_palette("FantasticFox1")) +
 theme(text = element_text("Georgia"))

fit_visual


## Regression Output

fit <- lm(constch_overall ~ corrupti + avg_yrsoffc, data = data)

modelsummary::modelsummary(fit,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("")) 








```


## Conclusion


To conclude, by confirming the positive correlation between the level of corruption and the number of constitutional changes, the analysis above affirms the hypothesis. However, by demonstrating that the coefficient of the mean Chief Executive term in the regression model is not statistically significant, the analysis also indicates the nuanced nature of the factors affecting the correlations. Therefore, it is important to acknowledge the limitations of the study, which, despite the rejection of the null hypothesis, do not allow us to assert with confidence a causal relationship between the level of corruption and the number of constitutional changes. Such limitations include the relatively short time-frame within which the data was collected (two decades), possible omitted variables and confounding variables, such as, for instance, the larger global factors and international pressures that may have affected constitutional changes. To alleviate some of these limitations would involve additional time as well as funds. If I were to improve my analysis using these additional resources, I would include data over the last century, rather than two decades, controlling for confounders, such as those involving historical context and taking into account the nature of the constitutional changes. Alas, that is beyond the scope of this study.
